createData_linear_modeling

### 2016-06-19 portfolio/analysis/6a-createData_linear_modeling_diff.Rmd
upload differenced model

@@

## Filtering out people with 10+ years of fishing. Drops rows from 13.7 to 10 million. 
```{r}
cfecnew = group_by(cfecnew, p_holder) %>% 
  mutate(nyr = length(unique(year))) %>% 
  filter(nyr >= 10)
```

@@

CUTOFF = 50
cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
```

@@

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg2[1], 
    days = diff(range(day) + 1), 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    length = mean(VLength, na.rm=T),
    strategy = strategy[1])

cfecAnnual = group_by(cfecAnnual, p_holder, year) %>%
  filter(npermit == 1)

@@

# random slope by p_holder
# Random intercepts in people, random intercepts in strategies, random slopes in strategies
# Group level predictors on intercepts 
# Indicator variable for EVOS years. Or whether

@@

#temp = group_by(cfecnew, p_holder, year, specn) %>% 
#  summarize(totRev.spec = sum(g_earn), 
#    permit = p_fshy[1])
#temp = left_join(temp, group_by(cfecnew, p_holder, year) %>% 
#    summarize(totRev = sum(g_earn)))
#temp$totRev.spec = temp$totRev.spec / temp$totRev

#mkt = group_by(cfecnew, p_fshy, year, specn) %>% 
#  summarize(totRev.mkt.spec = sum(g_earn)) 
#mkt = left_join(mkt, group_by(mkt, p_fshy, year) %>% 
#  summarize(totRev.mkt = sum(totRev.mkt.spec))) 
#mkt$totRev.mkt.spec = mkt$totRev.mkt.spec / mkt$totRev.mkt
#names(mkt)[1] = "permit"

# combine individual and market datasets
#temp = left_join(temp, mkt)

# Calculate squared proportions
#temp$p2 = (temp$totRev.spec - temp$totRev.mkt.spec)^2

#cfecAnnual = left_join(cfecAnnual, dplyr::select(temp, p_holder, year, p2))

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
#cfecAnnual$p2.prev = cfecAnnual$p2[idx]

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev
# idx = which(cfecAnnual$strategy != cfecAnnual$strategy.prev)

# Filter out salmon people
cfecAnnual$salmon = ifelse(substr(cfecAnnual$permit,1,1)=="S", "Y", "N")

cfecAnnual.salmon = cfecAnnual[cfecAnnual$salmon=="Y",]
cfecAnnual.nonsalmon = cfecAnnual[cfecAnnual$salmon=="N",]

#cfecAnnual.salmon = cfecAnnual.salmon[is.na(cfecAnnual.salmon$specDiv.change)==F, ]
#cfecAnnual.nonsalmon = cfecAnnual.nonsalmon[is.na(cfecAnnual.nonsalmon$specDiv.change)==F, ]

saveRDS(cfecAnnual.salmon, file="salmon_linearModeling_complete.rds")
saveRDS(cfecAnnual.nonsalmon, file="nonsalmon_linearModeling_complete.rds")

@@

## Do some exploration by permit types + diversity

For example, when someone is diversifying with a herring, halibut, or crab permit, what does that diversification look like?

```{r, echo=FALSE}
cfecnew$taxa = substr(cfecnew$p_fshy, 1, 1)
specPermit = group_by(cfecnew, taxa, specn) %>% 
  summarize(percent.revenue = sum(g_earn))
specPermit.tot = group_by(specPermit, taxa) %>% 
  summarize(tot = sum(percent.revenue))
specPermit = left_join(specPermit, specPermit.tot)
specPermit$percent.revenue = specPermit$percent.revenue/specPermit$tot
specPermit.df = group_by(specPermit, taxa, specn) %>% 
  filter(percent.revenue > 0.02) %>% 
  arrange(taxa) %>% 
  as.data.frame
specPermit.df$spec = cfec$spec[match(specPermit.df$specn, cfec$specn)]
kable(specPermit.df)
```

## Code for looking at permit 
We can look at whether there's clear combinations of permit types that are combinations -- most of these people either have a single local permit (PWS, Bristol Bay, etc) or have multiple statewide permits.

cfecnew$area = substr(cfecnew$p_fshy,4,4)
cfecnew$area[cfecnew$area %in% c("A","C")] = "Southeast"
cfecnew$area[cfecnew$area == "B"] = "Statewide"
cfecnew$area[cfecnew$area == "D"] = "Yakutat"
cfecnew$area[cfecnew$area == "E"] = "PWS"
cfecnew$area[cfecnew$area == "H"] = "CookInlet"
cfecnew$area[cfecnew$area == "J"] = "Westward"
cfecnew$area[cfecnew$area == "K"] = "Kodiak"
cfecnew$area[cfecnew$area == "L"] = "Chignik"
cfecnew$area[cfecnew$area == "M"] = "AlaskaPeninsula"
cfecnew$area[cfecnew$area == "N"] = "Nelson"
cfecnew$area[cfecnew$area == "Q"] = "BeringSea"
cfecnew$area[cfecnew$area == "T"] = "BristolBay"
cfecnew$area[cfecnew$area == "X"] = "Kotzebue"
cfecnew$area[cfecnew$area == "Z"] = "NortonSound"

cfecnew$spec = substr(cfecnew$p_fshy,1,1)
cfecnew$spec[cfecnew$spec == "B"] = "Halibut"
cfecnew$spec[cfecnew$spec == "C"] = "Sablefish"
cfecnew$spec[cfecnew$spec %in% c("D","E","K","T")] = "Crab"
cfecnew$spec[cfecnew$spec %in% c("F")] = "Freshwater"
cfecnew$spec[cfecnew$spec %in% c("G","H","L")] = "Herring"
cfecnew$spec[cfecnew$spec %in% c("I")] = "Lingcod"
cfecnew$spec[cfecnew$spec %in% c("A","J","N","R","W")] = "Shellfish"
cfecnew$spec[cfecnew$spec %in% c("O","P","Q","U","Z")] = "Invert"
cfecnew$spec[cfecnew$spec %in% c("M")] = "Misc"
cfecnew$spec[cfecnew$spec %in% c("S")] = "Salmon"
cfecnew$area.spec = paste(cfecnew$area, cfecnew$spec, sep="_")

cfecAnnual.multiplePermits = group_by(cfecnew, p_holder, year) %>%
  summarize(npermit = length(unique(p_fshy)), 
    allpermits = paste(sort(unique(p_fshy)), collapse = " "),
    areaspec = paste(sort(unique(area.spec)), collapse = " ")) %>%
  filter(npermit > 1)

>>>>

## Modifications from un-differenced filtering:  
### 1. No threshold for people to be included (e.g. 10+ years)
### 2. Inclusion of everyone, regardless of permits fished
### 3. Excluding observations that we don't have consecutive years for
### 4. Some permits lumped -- especially those based on vessel sizes

@@

## Aggregate some permits, especially some of the longline/trawl ones based on vessel size, for which reporting is also variable

```{r}
cfecnew$p_fshy[cfecnew$p_fshy=="B61B"] = "B06B"
cfecnew$p_fshy[cfecnew$p_fshy=="C61B"] = "C06B"
cfecnew$p_fshy[cfecnew$p_fshy=="C91B"] = "C09B"
cfecnew$p_fshy[cfecnew$p_fshy=="M61B"] = "M06B"
cfecnew$p_fshy[cfecnew$p_fshy=="E09B"] = "E91B"
cfecnew$p_fshy[cfecnew$p_fshy=="I06B"] = "I61B"
cfecnew$p_fshy[cfecnew$p_fshy=="Y06A"] = "Y61A"
cfecnew$p_fshy[cfecnew$p_fshy=="O06B"] = "O61B"
cfecnew$p_fshy[cfecnew$p_fshy=="O09B"] = "O91B"

cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FB","M7GB", "M7HB","M7IB")] = "M7IB"
cfecnew$p_fshy[cfecnew$p_fshy%in%c("M7FG","M7GG", "M7HG","M7IG")] = "M7IG"

# crab and shrimp vessels 
cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")] = paste("D09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="D91")])), sep="")
cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")] = paste("T09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="T91")])), sep="")
cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")] = paste("P09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="P91")])), sep="")

cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")] = paste("M09", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M91")])), sep="")
cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")] = paste("M06", substr(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")],4,length(cfecnew$p_fshy[which(substr(cfecnew$p_fshy,1,3)=="M61")])), sep="")
```

@@

#CUTOFF = 100
#cfecnew = cfecnew[cfecnew$strategy%in%names(rev(sort(table(topstrategies$strategy)))[1:CUTOFF]), ]
```

@@

cfecAnnual = group_by(cfecnew, p_holder, year) %>%
  summarize(permit = p_fshy[1], boat = cadfg2[1], 
    days = diff(range(day) + 1), 
    ngear = length(unique(gearn)),
    nspec = length(unique(specn)),
    npermit = length(unique(p_fshy)),
    revenue = sum(g_earn),
    length = mean(VLength, na.rm=T),
    strategy = strategy[1],
    weight = sum(g_pounds))

@@

@@

@@

cfecAnnual$permit.prev = cfecAnnual$permit[idx]
cfecAnnual$ngear.prev = cfecAnnual$ngear[idx]
cfecAnnual$nspec.prev = cfecAnnual$nspec[idx]
cfecAnnual$npermit.prev = cfecAnnual$npermit[idx]
cfecAnnual$revenue.prev = cfecAnnual$revenue[idx]
cfecAnnual$specdiv.prev = cfecAnnual$specDiv[idx]
cfecAnnual$geardiv.prev = cfecAnnual$gearDiv[idx]
cfecAnnual$strategy.prev = cfecAnnual$strategy[idx]
cfecAnnual$weight.prev = cfecAnnual$weight[idx]
cfecAnnual$days.prev = cfecAnnual$days[idx]
#cfecAnnual$p2.prev = cfecAnnual$p2[idx]

cfecAnnual$specDiv.change = cfecAnnual$specDiv - cfecAnnual$specdiv.prev

cfecAnnual.diff = cfecAnnual[is.na(cfecAnnual$specDiv.change)==F, ]

saveRDS(cfecAnnual.diff, file="diff_linearModeling_complete.rds")

# 4200 different strategies, need to model only most common, 
top.strategies = names(rev(sort(table(cfecAnnual.diff$strategy)))[1:100])

cfecAnnual.diff = cfecAnnual.diff[cfecAnnual.diff$strategy%in%top.strategies, ]
# idx = which(cfecAnnual.diff$strategy != cfecAnnual.diff$strategy.prev)
# paste(cfecAnnual.diff$strategy.prev[idx], cfecAnnual.diff$strategy[idx], sep=":")
cfecAnnual.diff = cfecAnnual.diff[which(cfecAnnual.diff$strategy == cfecAnnual.diff$strategy.prev), ]

saveRDS(cfecAnnual.diff, file="diff_linearModeling_subset.rds")

@@
